
---
title: "STAT 3240 Project Proposal"
author: "PROJECT GROUP"
date: "01/10/2024 `r Sys.Date()`"
output:
  html_document:
    fig_caption: yes
    theme: lumen
    toc: yes
    toc_depth: 2
    df_print: kable
    toc_float:
      collapsed: no
---

```{r, message=FALSE}
# Load all packages here:
library(readr)
library(dplyr)
library(ggplot2)
library(janitor)

# Set seed value of random number generator to get "replicable" random numbers.
# The choice of seed value of 76 was an arbitrary one on my part.
set.seed(76)
```

# Big-picture

## Research question

_What is your research question?_
How does the use of Artificial Intelligence affect college student grade and academic performance?

## Description of data

_Please give a very short description of the data set along with it's original source._
The data set, sourced from Kaggle, provides information on various factors influencing college students' academic performance, with a focus on the knowledge of Artificial Intelligence (AI). 


# Data

## Load data into R

_Include the code to load your data here. If your data is not confidential nor is it private in nature, consider publishing it as a `.csv` file on Google Sheets as in the code chunk below; instructions on how to do this are in Steps 1-6 [here](https://twitter.com/rudeboybert/status/1055821846539591680). If the data shouldn't be published online, then please submit the spreadsheet file on Moodle._

```{r}
library(readr)
studentp <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vRhKd_zI7lzSzqg6sG4aSes4GqGNKJr672S_MjGDuaQdXVXTl5GIQ6B-uSg9i_sD1vnzsuottRg0iy9/pub?gid=763522175&single=true&output=csv")
```

## Clean variable names

_Pipe your data frame into the `clean_names()` function from the `janitor` package. Then be sure to overwrite the original data frame with this new data frame using the `<-` assignment operator. This will clean your variable names, making them easier to work with._

```{r}
studentp <- studentp %>%
  clean_names()
glimpse(studentp)
```

## Data wrangling

_Complete your data wrangling here_: 

```{r}

```



***


# Preview of data

## Pare down variables

_`select()` the following variables **in this order** and drop all others. Eliminating all unnecessary variables will making visually exploring the raw values less taxing mentally, as we'll have less data to look at._

1. _First: The identification variable (if any)_
1. _Second: The outcome variable $y$_
1. _Third: The numerical explanatory variable_
1. _Fourth: The categorical explanatory variable_
1. _Rest: any other variables you find interesting_

```{r}
studentp <- studentp %>%
    select(writing_score, q1_ai_knowledge, q13_year_of_study , q16_gpa)

```

## Look at your data using glimpse

_Look at your data using the `glimpse()` function._

```{r}
glimpse(studentp)
```

## Show a preview of your data

_Look at your data another way by displaying a random sample of 5 rows of your data frame by piping it into the `sample_n(5)` function from the `dplyr` package._

```{r}
studentp %>%
  sample_n(5)
na.omit(studentp)

```



***



# Exploratory data analysis

_Let's do an little exploratory data analysis._

## Inspect for missing values

_Address missing values._


```{r}
colSums(is.na(studentp))
```


## Summary statistics

_Compute some quick summary statistics of the outcome variable and comment._

```{r}
studentp %>%
  group_by(q13_year_of_study) %>%
  summarize(n=n(),
            correlation = cor(writing_score, q1_ai_knowledge),
            mean = mean(writing_score, na.rm = TRUE),
            median = median(writing_score, na.rm = TRUE),
            sd = sd(writing_score, na.rm = TRUE))

```


## Histogram of outcome variable 

_Visualize the distribution of the outcome variable using a histogram and comment._

```{r, fig.cap = "Figure 1. Histogram of Student WWriting Score and Number of Students", fig.align = "center", fig.width=16/2, fig.height=9/2}
ggplot(studentp, aes( x= writing_score))+
  geom_histogram(binwidth= 2, color = "white", fill = "green")+
  labs(x = "Student Writing Score", y = "Number of Students")
```



## Scatterplot 

_Visualize the relationship of the outcome variable and the numerical explanatory variable using a scatterplot and comment._

```{r, fig.cap = "Figure 2. Scatterplot of Student Knowledge of AI against Student Writing Score", fig.align = "center", fig.width=16/2, fig.height=9/2}
ggplot(studentp, aes(x = q1_ai_knowledge, y = writing_score ))+
  geom_point() +
  geom_smooth(method = "lm", se = FALSE ) +
  labs(y = "Students Writing Score", 
       x = "Student Knowledge of AI")
```


## Boxplot

_Visualize the relationship of the outcome variable and the categorical explanatory variable using a scatterplot and comment._

```{r, fig.cap = "Figure 3. Boxplot of Years of Study and Student Writing Score", fig.align = "center", fig.width=16/2, fig.height=9/2}
ggplot(studentp, aes(x = q13_year_of_study, y = writing_score, fill = q13_year_of_study)) +
  geom_boxplot() +
  labs(y = "Student Writing Scores", x = "Years of Study")
```

## Colored scatterplot

_Visualize the relationship of the outcome variable and both explanatory variables using a colored scatterplot and comment._

```{r, fig.cap = "Figure 4. Scatterplot of Student AI Knowledge and Writing Scores", fig.align = "center", fig.width=16/2, fig.height=9/2}
ggplot(studentp, aes(x = q1_ai_knowledge, y = writing_score, color = q13_year_of_study))+
  geom_point() +
  geom_smooth(method = "lm", se = FALSE ) +
  labs(y = "Student Writing Score", x = "Student AI Knowledge")
```




# Author Statement

## Individual Roles

_Describe the role (Facilitator, Recorder, Reporter, Time Keeper, Checker) and level of contribution of each group member during this project phase_ 
Facilitator: Olamipo Amusa
Recorder: Parvesh Kumar Jogeea
Reporter: Arianna Lockhart
Time Keeper: Kevin Davies
Checker: Oluwatomilayo Faloseyi

## Individual Contribution

_If your group were to earn 5 points on this submission, how should those points be shared across your group members?_
Olampio Amusa = 1
Parvesh Kumar Jogeea = 1
Arianna Lockhart = 1
Kevin Davies = 1
Oluwatomilayo Faloseyi = 1

